---
title: "NFL_ggplot"
author: "Mark Paquette"
date: "2023-10-20"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data.set <- read.table("players.csv", header=TRUE, sep=",")
attach(data.set)
library(ggplot2)
```

# Graph One: Violin Plot

```{r, echo=FALSE, warning=FALSE}
ggplot(data.set,aes(x=position, y=weight)) + geom_violin(fill="red")
```

This is a violin plot and helps us visualize a numerical value (weight) vs
a categorical value (player position). This graph helps us gain an understanding
of both what the most common weights are at a given position while also showing
some of the outlier weight (for example a FB who weighs about 310 lbs). This
helps us answer questions like what is the typical weight for a Quarterback?

Since it is by organized by position, this type of graph does not really tell
us the statistics of the weight as a whole. To find this we would want a 
plot that only uses weight and not position

# Graph Two: Pie Chart

```{r, echo=FALSE, warning=FALSE}
pos <- c("C","CB","DB","DE","DT","FB","FS","G","ILB","LS","MLB","NT","OLB","QB",
         "RB","SS","T","TE","WR")
position_counts <- table(data.set$position)
position_counts <- position_counts[match(pos, names(position_counts))]

ggplot(data.frame(pos, position_counts), aes(x = "", y = position_counts, fill = pos)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void()  # Remove unnecessary elements
```

This Pie chart helps us visualize the number of players at each position. This
helps us gain an understanding what the most common and least common positions
are etc. One thing this is missing is exact numbers. We are not able to tell
how many of each position there are, we can only estimate how they compare to
one another. To actually represent this, we would need the numerical values 
for each position labeled on the chart or part of the legend.


